# Initial run - three cues, one set

This experiment is located in `experiments/exploratory/2022_07_28__three_cues_one_set`. 

The experiment aimed to see what behaviors evolved in an environment with three cues (with one being set to always be 1) and with naive parameters. 

## Dependencies

```{r, message=FALSE}
# External
library(ggplot2)
library(dplyr)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
# Internal
source('../global_shared_files/constant_vars__three_cues_one_set.R')
source('../global_shared_files/shared_funcs__three_cues_one_set.R')
```

## Data prep 

```{r}
df = read.csv('../experiments/exploration/2022_07_28__three_cues_one_set/data/combined_final_dominant_data.csv')
df = classify_individual_trials(df)
df = classify_seeds(df)
df_summary = summarize_final_dominant_org_data(df)
classification_summary = summarize_classifications(df_summary)
```

## Classification summary 

How many replicates evolved each behavior?

```{r, message=FALSE, warning=FALSE}
ggplot(classification_summary, aes(x = seed_classification_factor, y = count, fill = seed_classification_factor)) + 
  geom_col() + 
  geom_text(aes(y = count + 3, label = count)) + 
  scale_fill_manual(values = color_map) + 
  xlab('Classification') + 
  ylab('Number of replicates') + 
  theme(legend.position = 'none')
```

## Accuracy plots

Order the points by mean accuracy

```{r, message=FALSE, warning=FALSE}
df_summary = df_summary[order(df_summary$accuracy_mean),]
df_summary$seed_order = 1:nrow(df_summary)
df$seed_order = NA
for(seed in unique(df$seed)){
  df[df$seed == seed,]$seed_order = df_summary[df_summary$seed == seed,]$seed_order
}
```

For each replicate, plot mean accuracy as a solid point and min/max as partially transparent points.

```{r, message=FALSE, warning=FALSE}
ggplot(df_summary, aes(x = seed_order, color = seed_classification)) + 
  geom_point(aes(y = accuracy_mean)) +
  geom_point(aes(y = accuracy_min), alpha = 0.2) +
  geom_point(aes(y = accuracy_max), alpha = 0.2) +
  scale_color_manual(values = color_map) +
  xlab('Replicates (ordered)') +
  ylab('Accuracy') +
  labs(color = 'Classification')
```

Now plot the accuracy of each sample, maintaining the same order on the x-axis. 

```{r, message=FALSE, warning=FALSE}
ggplot(df, aes(x = seed_order, y = accuracy, color = seed_classification)) + 
  geom_point(alpha = 0.2) + 
  scale_color_manual(values = color_map) +
  xlab('Replicates (ordered)') +
  ylab('Accuracy') +
  labs(color = 'Classification') 
```

## Raincloud plots (individual trials)

```{r, message=FALSE, warning=FALSE}
# Raincloud plot of accuracy
ggplot(df, aes(x = seed_classification_factor, y = accuracy, fill = seed_classification_factor)) + 
  geom_flat_violin(scale="width", position = position_nudge(x = .2, y = 0), alpha = .8 ) + 
  geom_point(mapping=aes(color=seed_classification_factor), position = position_jitter(width = .15, height = 0), size = .5, alpha = 0.8 ) + 
  geom_boxplot( width = .1, outlier.shape = NA, alpha = 0.5 ) +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) + 
  xlab('Classification') + 
  ylab('Accuracy') + 
  theme(legend.position = 'none')
```

```{r, message=FALSE, warning=FALSE}
# Raincloud plot of merit
ggplot(df, aes(x = seed_classification_factor, y = merit, fill = seed_classification_factor)) + 
  geom_flat_violin(scale="width", position = position_nudge(x = .2, y = 0), alpha = .8 ) + 
  geom_point(mapping=aes(color=seed_classification_factor), position = position_jitter(width = .15, height = 0), size = .5, alpha = 0.8 ) + 
  geom_boxplot( width = .1, outlier.shape = NA, alpha = 0.5 ) +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) + 
  xlab('Classification') + 
  ylab('Merit') + 
  theme(legend.position = 'none')
```

## Raincloud plots (per replicate)
```{r, message=FALSE, warning=FALSE}
# Raincloud plot of genome length
ggplot(df_summary, aes(x = seed_classification_factor, y = genome_length, fill = seed_classification_factor)) + 
  geom_flat_violin(scale="width", position = position_nudge(x = .2, y = 0), alpha = .8 ) + 
  geom_point(mapping=aes(color=seed_classification_factor), position = position_jitter(width = .15, height = 0), size = .5, alpha = 0.8 ) + 
  geom_boxplot( width = .1, outlier.shape = NA, alpha = 0.5 ) +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) + 
  xlab('Classification') + 
  ylab('Genome length') + 
  theme(legend.position = 'none')
```

```{r, message=FALSE, warning=FALSE}
# Raincloud plot of mean accuracy of replicates
ggplot(df_summary, aes(x = seed_classification_factor, y = accuracy_mean, fill = seed_classification_factor)) + 
  geom_flat_violin(scale="width", position = position_nudge(x = .2, y = 0), alpha = .8 ) + 
  geom_point(mapping=aes(color=seed_classification_factor), position = position_jitter(width = .15, height = 0), size = .5, alpha = 0.8 ) + 
  geom_boxplot( width = .1, outlier.shape = NA, alpha = 0.5 ) +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) + 
  scale_y_continuous(limits = c(0,1)) +
  xlab('Classification') + 
  ylab('Mean accuracy') + 
  theme(legend.position = 'none')
```

```{r, message=FALSE, warning=FALSE}
# Raincloud plot of mean merit of replicates
ggplot(df_summary, aes(x = seed_classification_factor, y = merit_mean, fill = seed_classification_factor)) + 
  geom_flat_violin(scale="width", position = position_nudge(x = .2, y = 0), alpha = .8 ) + 
  geom_point(mapping=aes(color=seed_classification_factor), position = position_jitter(width = .15, height = 0), size = .5, alpha = 0.8 ) + 
  geom_boxplot( width = .1, outlier.shape = NA, alpha = 0.5 ) +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) + 
  xlab('Classification') + 
  ylab('Mean merit') + 
  theme(legend.position = 'none')
```
